\ourlongtitle

### 2021-08-28 Bernoulli/glo-2020-numerics.tex
Final correction (typos)

In order to further illustrate the $\mathcal{O}(n \precpar)$ bound provided by Theorem~\ref{thm:bias:bound} as well as the stochastic stability of Algorithm~\ref{alg:ideal:PaRIS} established in Corollary~\ref{cor:variance:bound:ideal:PaRIS}, Kalman smoothing was conducted for $n \in \intvect{1}{50}$ on the basis of a skew model with fixed $\varepsilon = 0.1$. Figure~\ref{fig:varying:n} shows, as expected, a linear increase of the bias with $n$ (red line). In addition, displaying also the errors (green lines) of $60$ independent ideal PaRIS replicates (obtained under the skew model with the same algorithmic parametrisation as previously), the same plot confirms the stochastic stability of the ideal PaRIS algorithm, as the variance does not grow faster than linearly with $n$. Finally, for completeness Figure~\ref{fig:varying:n} reports similar errors (blue lines) when the ideal PaRIS evolves under the dynamics of the true model, and in this case the bias is negligible as expected.% (whereas the variance is still increasing linearly).

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{%OL-version/figures/
        vary_epsilon_fixed_n.png}
        \caption{}
        \label{fig:varying:precpar}
    \end{subfigure}
    ~ 
    \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=\textwidth]{%OL-version/figures/
        fixed_epsilon_vary_n.png}
        \caption{}
        \label{fig:varying:n}
    \end{subfigure}
    \caption{(a) The red line is the deviation of $\postmod{0:n} h_n$ from $\post{0:n} h_n$ for varying $\precpar \in \{0, 0.05, 0.1, \ldots, 0.5 \}$ and fixed $n = 50$, computed by means of Kalman smoothing. Turquoise markers are similar errors obtained on the basis of $60$ replicates of the ideal PaRIS in Algorithm~\ref{alg:ideal:PaRIS} with $\N = 200$ particles and $\K = 2$ backward samples. (b) The red line is the same deviation for increasing $n \in \intvect{1}{50}$ and fixed $\precpar = 0.1$. Green and blue lines correspond to errors of $60$ independent ideal PaRIS replicates obtained under the skew and true models, respectively, for $\N = 200$ and $\K = 2$.} \label{fig}
\end{figure}

>>>>

In order to further illustrate the $\mathcal{O}(n \precpar)$ bound provided by Theorem~\ref{thm:bias:bound} as well as the stochastic stability of Algorithm~\ref{alg:ideal:PaRIS} established in Corollary~\ref{cor:variance:bound:ideal:PaRIS}, Kalman smoothing was conducted for $n \in \intvect{1}{50}$ on the basis of a skew model with fixed $\varepsilon = 0.1$. Figure~\ref{fig:varying:n} shows, as expected, a linear increase of the bias with $n$ (red line). In addition, displaying also the errors (green lines) of $60$ independent ideal PaRIS replicates (obtained under the skew model), the same plot confirms the stochastic stability of the ideal PaRIS algorithm, as the variance does not grow faster than linearly with $n$. Finally, for completeness Figure~\ref{fig:varying:n} reports similar errors (blue lines) when the ideal PaRIS evolves under the dynamics of the true model, and in this case the bias is negligible as expected.% (whereas the variance is still increasing linearly).

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.43\textwidth}
        \includegraphics[width=\textwidth]{%OL-version/figures/
        vary_epsilon_fixed_n.png}
        \caption{}
        \label{fig:varying:precpar}
    \end{subfigure}
    ~ 
    \begin{subfigure}[b]{0.43\textwidth}
        \includegraphics[width=\textwidth]{%OL-version/figures/
        fixed_epsilon_vary_n.png}
        \caption{}
        \label{fig:varying:n}
    \end{subfigure}
    \caption{(a) The red line is the deviation of $\postmod{0:n} h_n$ from $\post{0:n} h_n$ for varying $\precpar \in \{0, 0.05, 0.1, \ldots, 0.5 \}$ and fixed $n = 50$, computed by means of Kalman smoothing. Turquoise markers are similar errors obtained on the basis of $60$ replicates of the ideal PaRIS in Algorithm~\ref{alg:ideal:PaRIS} with $\N = 200$ particles and $\K = 2$ backward samples. (b) The red line is the same deviation for increasing $n \in \intvect{1}{50}$ and fixed $\precpar = 0.1$. Green and blue lines correspond to errors of $60$ independent ideal PaRIS replicates obtained under the skew and true models, respectively, for $\N = 200$ and $\K = 2$.} \label{fig}
\end{figure}

### 2021-08-28 Bernoulli/glo-2020-pm-PaRIS.tex
Final correction (typos)

\begin{hypH}
\label{assum:biased:estimate}
Let $(\Zset_n, \Zfd_n)_{n \in \nsetpos}$ be a sequence of general state spaces. For each $n \in \nset$ there exist a Markov kernel $\ukdist{n}$ on $\Xset_n \times \Xset_{n + 1} \times \Zfd_{n + 1}$ and a positive measurable function $\ukest{n}{z}(x_n, x_{n + 1})$ on $\Xset_n \times \Xset_{n + 1} \times \Zset_{n + 1}$ such that for every $x_{n:n + 1} \in \Xset_n \times \Xset_{n + 1}$, drawing $\zeta \sim \ukdist{n}(x_{n:n + 1}, \rmd z)$ and computing $\ukest{n}{\zeta}(x_n, x_{n + 1})$ yields an estimate of $\ud{n}(x_n, x_{n + 1})$.  
\end{hypH}

Under \hypref{assum:biased:estimate}, we denote, for every $n \in \nset$ and $x_{n:n + 1} \in \Xset_n \times \Xset_{n + 1}$, by   
\begin{equation} \label{eq:def:udmod}
\udmod{n}(x_n, x_{n + 1}) \eqdef \int \ukdist{n}(x_{n:n + 1}, \rmd z) \, \ukest{n}{z}(x_n, x_{n + 1})
\end{equation}
the expectation of the statistic $\ukest{n}{\zeta}(x_n, x_{n + 1})$. Here $\precpar$ is an accuracy parameter belonging to some parameter space $\precparsp \subset \rset$ and controlling the bias of the estimated model with respect to the true model; this will be discussed in depth in Section~\ref{sec:lipschitz:continuity}. For each $n \in \nset$ we define the unnormalised kernel 
\begin{equation} \label{eq:def:ukmod}
    \ukmod{n}(x_n, \rmd x_{n + 1}) = \udmod{n}(x_n, x_{n + 1}) \, \mu_{n + 1}(\rmd x_{n + 1}) 
\end{equation}
on $\Xset_n \times \Xfd_{n + 1}$.

@@

\begin{algorithm}[h] 
    \KwData{$(\epart{n}{i}, \ewght{n}{i})_{i = 1}^\N$}
    \KwResult{$(\epart{n + 1}{i}, \ewght{n + 1}{i})_{i = 1}^\N$}
    \For {$i = 1 \to \N$}{
        draw $\ind{n + 1}{i} \sim \cat(\{ \adjfuncforward{n}(\epart{n}{\ell}) \ewght{n}{\ell} \}_{\ell = 1}^\N)$\;
        draw $\epart{n + 1}{i} \sim \prop{n}(\epart{n}{{\ind{n + 1}{i}}}, \cdot)$\;
        draw $\zpart{n + 1}{i} \sim \ukdist{n}(\epart{n}{{\ind{n + 1}{i}}}, \epart{n + 1}{i}, \cdot)$\;
        set $\displaystyle 
        \ewght{n + 1}{i} \gets \frac{\ukest{n}{\zpart{n + 1}{i}}(\epart{n}{{\ind{n + 1}{i}}}, \epart{n + 1}{i})}{\adjfuncforward{n}(\epart{n}{{\ind{n + 1}{i}}}) \propdens{n}(\epart{n}{{\ind{n + 1}{i}}}, \epart{n + 1}{i})}$\;
}
\caption{Pseudo-marginal forward sampling, \textsf{pmFS}.} \label{alg:pm:SMC}
\end{algorithm}

Iterating recursively, after initialisation as in Section~\ref{sec:SMC}, pseudo-marginal forward sampling yields a generalisation of the random-weight particle filter proposed in \cite{fearnhead2008particle} in the context of partially observed diffusion processes.

>>>>

\begin{hypH}
\label{assum:biased:estimate}
Let $(\Zset_n, \Zfd_n)_{n \in \nsetpos}$ be a sequence of general state spaces. For each $n \in \nset$ there exist a Markov kernel $\ukdist{n}$ on $\Xset_n \times \Xset_{n + 1} \times \Zfd_{n + 1}$ and a positive measurable function $\ukest{n}{z}(x_n, x_{n + 1})$ on $\Xset_n \times \Xset_{n + 1} \times \Zset_{n + 1}$ such that for every $x_{n:n + 1} \in \Xset_n \times \Xset_{n + 1}$, drawing $\zeta \sim \ukdist{n}(x_{n:n + 1}, \rmd z)$ and computing $\ukest{n}{\zeta}(x_n, x_{n + 1})$ yields an estimate of $\ud{n}(x_n, x_{n + 1})$.  
\end{hypH}
Under \hypref{assum:biased:estimate}, we denote, for every $n \in \nset$ and $x_{n:n + 1} \in \Xset_n \times \Xset_{n + 1}$, by   
\begin{equation} \label{eq:def:udmod}
\udmod{n}(x_n, x_{n + 1}) \eqdef \int \ukdist{n}(x_{n:n + 1}, \rmd z) \, \ukest{n}{z}(x_n, x_{n + 1})
\end{equation}
the expectation of the statistic $\ukest{n}{\zeta}(x_n, x_{n + 1})$. Here $\precpar$ is an accuracy parameter belonging to some parameter space $\precparsp \subset \rset$ and controlling the bias of the estimated model with respect to the true model; this will be discussed in depth in Section~\ref{sec:lipschitz:continuity}. For each $n \in \nset$ we define the unnormalised kernel 
\begin{equation} \label{eq:def:ukmod}
    \ukmod{n}(x_n, \rmd x_{n + 1}) = \udmod{n}(x_n, x_{n + 1}) \, \mu_{n + 1}(\rmd x_{n + 1}) 
\end{equation}
on $\Xset_n \times \Xfd_{n + 1}$.

@@

\begin{algorithm}[h] 
    \KwData{$(\epart{n}{i}, \ewght{n}{i})_{i = 1}^\N$}
    \KwResult{$(\epart{n + 1}{i}, \ewght{n + 1}{i})_{i = 1}^\N$}
    \For {$i = 1 \to \N$}{
        draw $\ind{n + 1}{i} \sim \cat(\{ \adjfuncforward{n}(\epart{n}{\ell}) \ewght{n}{\ell} \}_{\ell = 1}^\N)$\;
        draw $\epart{n + 1}{i} \sim \prop{n}(\epart{n}{{\ind{n + 1}{i}}}, \cdot)$\;
        draw $\zpart{n + 1}{i} \sim \ukdist{n}(\epart{n}{{\ind{n + 1}{i}}}, \epart{n + 1}{i}, \cdot)$\;
        set $\displaystyle 
        \ewght{n + 1}{i} \gets \frac{\ukest{n}{\zpart{n + 1}{i}}(\epart{n}{{\ind{n + 1}{i}}}, \epart{n + 1}{i})}{\adjfuncforward{n}(\epart{n}{{\ind{n + 1}{i}}}) \propdens{n}(\epart{n}{{\ind{n + 1}{i}}}, \epart{n + 1}{i})}$\;
}
\caption{Pseudo-marginal forward sampling, \textsf{pmFS}.} \label{alg:pm:SMC}
\end{algorithm}
Iterating recursively, after initialisation as in Section~\ref{sec:SMC}, pseudo-marginal forward sampling yields a generalisation of the random-weight particle filter proposed in \cite{fearnhead2008particle} in the context of partially observed diffusion processes.

### 2021-08-28 Bernoulli/glo-2020-theor-results.tex
Final correction (typos)

The previous bound describes the pointwise convergence of the estimator delivered by Algorithm~\ref{alg:pm:PaRIS} as $\N$ grows for $n$ fixed, and here no attempt has been made to obtain a bound that is uniform in $n$. As we shall see shortly, in Theorem~\ref{thm:variance:bound}, the numerical stability of the algorithm can instead be established by bounding the asymptotic variance of the estimator. Finally, note that the previous bound implies that the estimator $\sum_{i = 1}^\N \ewght{n}{i} \tstat[i]{n} / \sumwght{n}$ tends $\pP$-a.s. to $\postmod{0:n} h_n$ as $\N$ tends to infinity.

@@

We now focus on the asymptotic properties of Algorithm~\ref{alg:pm:PaRIS} and furnish, in Theorem~\ref{cor:clt:pseudo:marginal:paris} below, this scheme with a CLT. On the basis of this result, the stochastic stability of the algorithm is expressed by establishing, in Theorem~\ref{thm:variance:bound}, an $\mathcal{O}(n)$ bound on the asymptotic variances. In order to be able to state these results accurately, we need some additional notation. First, let 
 for $(x_n, x_{n + 1}) \in \Xset_n \times \Xset_{n + 1}$, 
\begin{multline} \label{eq:def:ukestvar}
\ukestvar{n}(x_n, x_{n + 1}) \eqdef 
\frac{1}{\wgtfuncmod{n}(x_n, x_{n + 1})} \int \{ \wgtfunc{n}(x_n, x_{n + 1}, z) - \wgtfuncmod{n}(x_n, x_{n + 1}) \}^2 \, \ukdist{n}(x_n, x_{n + 1}, \rmd z)
\end{multline}
denote the conditional relative variance of the (random) weight $\ewght{n + 1}{i}$ given $\epart{n}{\ind{n}{i}} = x_n$ and $\epart{n + 1}{i}= x_{n + 1}$. In addition, for each $n \in \nset$ and $m \in \intvect{0}{n}$, define the kernel  
\begin{equation} \label{eq:def:retrokmodmod}
    \retrokmodmod_{m, n}(x_m', \rmd x_{0:n}) \eqdef \delta_{x_m'}(\rmd x_m) \,  
    \tstatmod{m}(x_m, \rmd x_{0:m - 1})
    \prod_{\ell = m}^{n - 1} \ukmod{\ell}(x_\ell, \rmd x_{\ell + 1})
\end{equation}
on $\Xset_m \times \Xfd^n$ as well as the centered version 
\begin{equation} \label{eq:def:retrokmodmodnorm}
\retrokmodmodnorm_{m, n} h (x_m) \eqdef  \retrokmodmod_{m, n}(h - \postmod{0:n} h)(x_m) 
\end{equation}
on the same space. The following is the main result of this section. 
\begin{theorem} \label{cor:clt:pseudo:marginal:paris}
Assume~\hypref[assum:biased:estimate]{assum:bound:filter:pseudomarginal}. Then for all $n \in \nset$, $\precpar \in \precparsp$, $\K \in \nsetpos$, and $h_n \in \bmaf{\Xfd^n}$, 
$$
 \sqrt{\N} \left( \sum_{i = 1}^\N \frac{\ewght{n}{i}}{\sumwght{n}} \tstat[i]{n} - \postmod{0:n} h_n  \right) 
  \dlim \sigma_n(h_n) Z, 
$$
where $Z$ is standard normally distributed and 
\begin{equation} \label{eq:as:var:decomp}
\sigma^2_n (h_n) \eqdef \frac{\chi(\initwgtfunc \retrokmodmodnorm_{0, n} h_n)^2}{(\chi \ukmod{0, n - 1} \1_{\Xset^n})^2} + \sigma^2_n \langle (\wgtfuncmod{\ell})_{\ell = 0}^{n - 1} \rangle (h_n) + \sigma^2_n \langle (\ukestvar{\ell})_{\ell = 0}^{n - 1} \rangle (h_n) 
\end{equation}
and
\begin{multline} \label{eq:non-recursive:as:var}
\sigma^2_n \langle (\varphi_\ell)_{\ell = 0}^{n - 1} \rangle (h_n) 
\eqdef \sum_{m = 0}^{n - 1} \frac{\postmod{m} \adjfuncforward{m} \postmod{m} \ukmod{m}(\varphi_m [\retrokmodmodnorm_{m + 1, n} h_n ]^2)}{(\postmod{m} \ukmod{m, n - 1} \1_{\Xset^n})^2}\\
+ \sum_{m = 0}^{n - 1} \sum_{\ell = 0}^m \frac{\postmod{m} \adjfuncforward{m} \postmod{\ell} \ukmod{\ell} \{\bkmod{\ell}(\tstatmod{\ell} h_{\ell} + \addf{\ell} - \tstatmod{\ell + 1} h_{\ell +1})^2 \ukmod{\ell + 1, m}( \bkmod{m} \varphi_m [\ukmod{m + 1, n - 1} \1_{\Xset^n}]^2
)\}}{\K^{m - \ell + 1} (\postmod{\ell} \ukmod{\ell, m - 1} \1_{\Xset^m})(\postmod{m} \ukmod{m, n - 1} \1_{\Xset^n})^2}
\end{multline}
for any sequence $(\varphi_\ell)_{\ell \in \nset}$ of measurable functions $\varphi_\ell : \Xset_\ell \times \Xset_{\ell + 1} \to \rsetnn$. 
\end{theorem}

@@

Having introduced these quantities, we are ready to present the main result of this section.

@@

\begin{corollary} \label{cor:variance:bound}
Assume \hypref[assum:biased:estimate]{assum:strong:mixing}. Then for every $\precpar \in \precparsp$, $\K \geq 2$, and $h_n \in \bmaf{\Xfd^n}$, 
\begin{multline} \label{eq:variance:bound:full:monty}
\limsup_{n \to \infty} \frac{1}{n} \sigma^2_n(h_n) \\ 
\leq \left( \cbound + \frac{1}{\K - 1} \dbound \right) \sup_{\ell \in \nset} \| \addf{\ell} \|_\infty^2 \sup_{\ell \in \nset} \| \adjfuncforward{\ell} \|_\infty  \left( \sup_{\ell \in \nset} \| \wgtfuncmod{\ell} \|_\infty + \sup_{\ell \in \nset} \| \ukestvar{\ell} \|_\infty \right) \\
+ \frac{1}{\rho^2(1 - \rho)^4 (\chi \1_{\Xset_0})^2} \sup_{\ell \in \nset} \| \addf{\ell} \|_\infty^2. 
\end{multline}
\end{corollary}

@@

The second case can be treated straightforwardly by combining the bound \eqref{eq:Euler:bound}, applied to the bivariate process $(X_t, Y_t)_{t > 0}$, with \eqref{eq:ukmod:durham:gallant:case:2}. This provides the existence of constants $\tilde{c}_\delta > 0$ and $\tilde{d}_\delta > 0$ such that for all $\precpar \in \precparsp_\delta$ and $(x_n, y_n, y_{n + 1}) \in \Xset \times \Yset^2$ and $h \in \bmf{\Xfd^{\tensprod 2}}$,  
\begin{align*}
\lefteqn{\left|\ukmod{n} h (x_n) - \uk{n} h (x_n) \right|}\hspace{20mm} \\
&\leq \tilde{c}_\delta \frac{\precpar}{\delta} \exp \left( - \tilde{d}_\delta \| y_{n + 1} -y_n \|^2\right) \int h(x_n, x_{n + 1}) \exp \left( - \tilde{d}_\delta \| x_{n + 1} - x_n \|^2\right) \, \rmd x_{n + 1} \\
&\leq \tilde{c}_\delta \frac{\precpar}{\delta} \left( \frac{\pi}{\tilde{d}_\delta} \right)^{d_x / 2} \|h \|_\infty.  
\end{align*}

@@

\begin{remark}
Consider now a parameterised version of the model, where the transition densities $(\ud{n ; \theta})_{k \in \nset}$ are indexed by some parameter $\theta$ belonging to some parameter space $\Theta$ being a subset of $\rset^d$. Assume further that all $(\ud{n ; \theta})_{n \in \nset}$ are differentiable with respect to $\theta$ and such that for all $n \in \nset$, $x_n \in \Xset$, and $h \in \bmf{\Xfd_{n + 1}}$,  
$$
\nabla_\theta \int \ud{n ; \theta}(x_n, x_{n + 1}) h(x_{n + 1}) \, \mu_{n + 1}(\rmd x_{n + 1}) = \int \nabla_\theta \ud{n ; \theta}(x_n, x_{n + 1}) h(x_{n + 1}) \, \mu_{n + 1}(\rmd x_{n + 1})
$$
and 
$$
\sup_{\theta \in \Theta} \int |\nabla_\theta \ud{n ; \theta}(x_n, x_{n + 1})| \, \mu_{n + 1}(\rmd x_{n + 1}) \leq c < \infty 
$$
for some positive constant $c$, implying \hypref{assum:bias:bound} in the sense that for all $n \in \nset$, $h \in \bmf{\Xfd_{n + 1}}$, and $x_n \in \Xset$, 
$$
| \uk{n; \theta}h(x_n) - \uk{n; \theta'}h(x_n) | \leq c \| \theta - \theta' \| \| h \|_\infty
$$
(\emph{i.e.}, $\precpar = \| \theta - \theta' \|$ in this case). Finally, assume also that family satisfies \hypref{assum:strong:mixing} uniformly over the parameter space in the sense that for all $n \in \nset$, $(x_n, x_{n + 1}) \in \Xset^2$, and $\theta \in \Theta$, 
$
\udlow \leq \ud{n; \theta}(x_n, x_{n + 1}) \leq \udup. 
$  
Then Theorem~\ref{thm:bias:bound} provides a positive constant $d$ such that for all $n \in \nset$,  
$$
\big| \post{0:n ; \theta} h_n -  \post{0:n ; \theta'} h_n \big|  
\leq d n \| \theta - \theta' \| \sup_{k \in \nset} \| \addf{k} \|_\infty. 
$$
This extends previous results on the uniform continuity of the filter distribution (see, \eg, \cite{papavasiliou:2006,legland:oudjane:2004}) to smoothing of additive state functionals.  
\end{remark}

>>>>

The previous bound describes the pointwise convergence of the estimator delivered by Algorithm~\ref{alg:pm:PaRIS} as $\N$ grows for $n$ fixed, and here no attempt has been made to obtain a bound that is uniform in $n$. As we shall see shortly, in Theorem~\ref{thm:variance:bound}, the numerical stability of the algorithm can instead be established by bounding the asymptotic variance of the estimator. %Finally, note that the previous bound implies that the estimator $\sum_{i = 1}^\N \ewght{n}{i} \tstat[i]{n} / \sumwght{n}$ tends $\pP$-a.s. to $\postmod{0:n} h_n$ as $\N$ tends to infinity.

@@

We now focus on the asymptotic properties of Algorithm~\ref{alg:pm:PaRIS} and furnish, in Theorem~\ref{cor:clt:pseudo:marginal:paris} below, this scheme with a CLT. The stochastic stability of the algorithm is expressed by establishing, in Theorem~\ref{thm:variance:bound}, an $\mathcal{O}(n)$ bound on the asymptotic variances. %In order to be able to state these results accurately, we need some additional notation. 
First, let for $(x_n, x_{n + 1}) \in \Xset_n \times \Xset_{n + 1}$, 
\begin{multline} \label{eq:def:ukestvar}
\ukestvar{n}(x_n, x_{n + 1}) \eqdef 
\frac{1}{\wgtfuncmod{n}(x_n, x_{n + 1})} \int \{ \wgtfunc{n}(x_n, x_{n + 1}, z) - \wgtfuncmod{n}(x_n, x_{n + 1}) \}^2 \, \ukdist{n}(x_n, x_{n + 1}, \rmd z)
\end{multline}
denote the conditional relative variance of the (random) weight $\ewght{n + 1}{i}$ given $\epart{n}{\ind{n}{i}} = x_n$ and $\epart{n + 1}{i}= x_{n + 1}$. In addition, for each $n \in \nset$ and $m \in \intvect{0}{n}$, define the kernel  
\begin{equation} \label{eq:def:retrokmodmod}
    \retrokmodmod_{m, n}(x_m', \rmd x_{0:n}) \eqdef \delta_{x_m'}(\rmd x_m) \,  
    \tstatmod{m}(x_m, \rmd x_{0:m - 1})
    \prod_{\ell = m}^{n - 1} \ukmod{\ell}(x_\ell, \rmd x_{\ell + 1})
\end{equation}
on $\Xset_m \times \Xfd^n$ as well as the centered version 
\begin{equation} \label{eq:def:retrokmodmodnorm}
\retrokmodmodnorm_{m, n} h (x_m) \eqdef  \retrokmodmod_{m, n}(h - \postmod{0:n} h)(x_m) 
\end{equation}
on the same space. The following is the main result of this section. 
\begin{theorem} \label{cor:clt:pseudo:marginal:paris}
Assume~\hypref[assum:biased:estimate]{assum:bound:filter:pseudomarginal}. Then for all $n \in \nset$, $\precpar \in \precparsp$, $\K \in \nsetpos$, and $h_n \in \bmaf{\Xfd^n}$, 
$$
 \sqrt{\N} \left( \sum_{i = 1}^\N \frac{\ewght{n}{i}}{\sumwght{n}} \tstat[i]{n} - \postmod{0:n} h_n  \right) 
  \dlim \sigma_n(h_n) Z, 
$$
where $Z$ is standard normally distributed and 
\begin{equation} \label{eq:as:var:decomp}
\sigma^2_n (h_n) \eqdef \frac{\chi(\initwgtfunc \retrokmodmodnorm_{0, n} h_n)^2}{(\chi \ukmod{0, n - 1} \1_{\Xset^n})^2} + \sigma^2_n \langle (\wgtfuncmod{\ell})_{\ell = 0}^{n - 1} \rangle (h_n) + \sigma^2_n \langle (\ukestvar{\ell})_{\ell = 0}^{n - 1} \rangle (h_n) 
\end{equation}
and
\begin{multline} \label{eq:non-recursive:as:var}
\sigma^2_n \langle (\varphi_\ell)_{\ell = 0}^{n - 1} \rangle (h_n) 
\eqdef \sum_{m = 0}^{n - 1} \frac{\postmod{m} \adjfuncforward{m} \postmod{m} \ukmod{m}(\varphi_m [\retrokmodmodnorm_{m + 1, n} h_n ]^2)}{(\postmod{m} \ukmod{m, n - 1} \1_{\Xset^n})^2}\\
+ \sum_{m = 0}^{n - 1} \sum_{\ell = 0}^m \frac{\postmod{m} \adjfuncforward{m} \postmod{\ell} \ukmod{\ell} \{\bkmod{\ell}(\tstatmod{\ell} h_{\ell} + \addf{\ell} - \tstatmod{\ell + 1} h_{\ell +1})^2 \ukmod{\ell + 1, m}( \bkmod{m} \varphi_m [\ukmod{m + 1, n - 1} \1_{\Xset^n}]^2
)\}}{\K^{m - \ell + 1} (\postmod{\ell} \ukmod{\ell, m - 1} \1_{\Xset^m})(\postmod{m} \ukmod{m, n - 1} \1_{\Xset^n})^2}
\end{multline}
for any sequence $(\varphi_\ell)_{\ell \in \nset}$ of measurable functions $\varphi_\ell : \Xset_\ell \times \Xset_{\ell + 1} \to \rsetnn$. 
\end{theorem}

@@

%Having introduced these quantities, we are ready to present the main result of this section.

@@

\begin{corollary} \label{cor:variance:bound}
Assume \hypref[assum:biased:estimate]{assum:strong:mixing}. Then for every $\precpar \in \precparsp$, $\K \geq 2$, and $h_n \in \bmaf{\Xfd^n}$, 
\begin{multline} \label{eq:variance:bound:full:monty}
\limsup_{n \to \infty} \frac{1}{n} \sigma^2_n(h_n) 
\leq \left( \cbound + \frac{1}{\K - 1} \dbound \right) \sup_{\ell \in \nset} \| \addf{\ell} \|_\infty^2 \sup_{\ell \in \nset} \| \adjfuncforward{\ell} \|_\infty  \left( \sup_{\ell \in \nset} \| \wgtfuncmod{\ell} \|_\infty + \sup_{\ell \in \nset} \| \ukestvar{\ell} \|_\infty \right) \\
+ \frac{1}{\rho^2(1 - \rho)^4 (\chi \1_{\Xset_0})^2} \sup_{\ell \in \nset} \| \addf{\ell} \|_\infty^2. 
\end{multline}
\end{corollary}

@@

The second case can be treated straightforwardly by combining the bound \eqref{eq:Euler:bound}, applied to the bivariate process $(X_t, Y_t)_{t > 0}$, with \eqref{eq:ukmod:durham:gallant:case:2}. This provides the existence of constants $\tilde{c}_\delta > 0$ and $\tilde{d}_\delta > 0$ such that for all $\precpar \in \precparsp_\delta$ and $(x_n, y_n, y_{n + 1}) \in \Xset \times \Yset^2$ and $h \in \bmf{\Xfd^{\tensprod 2}}$,  
\begin{align*}
\left|\ukmod{n} h (x_n) - \uk{n} h (x_n) \right| &\leq \tilde{c}_\delta \frac{\precpar}{\delta} \mathrm{e}^{- \tilde{d}_\delta \| y_{n + 1} -y_n \|^2} \int h(x_n, x_{n + 1}) \exp \left( - \tilde{d}_\delta \| x_{n + 1} - x_n \|^2\right) \, \rmd x_{n + 1} \\
&\leq \tilde{c}_\delta \frac{\precpar}{\delta} \left( \frac{\pi}{\tilde{d}_\delta} \right)^{d_x / 2} \|h \|_\infty.  
\end{align*}

@@

\begin{remark}
Consider now a parameterised version of the model, where the transition densities $(\ud{n ; \theta})_{k \in \nset}$ are indexed by some parameter $\theta$ belonging to some parameter space $\Theta$ being a subset of $\rset^d$. Assume further that all $(\ud{n ; \theta})_{n \in \nset}$ are differentiable with respect to $\theta$ and such that for all $n \in \nset$, $x_n \in \Xset$, and $h \in \bmf{\Xfd_{n + 1}}$,  
$$
\nabla_\theta \int \ud{n ; \theta}(x_n, x_{n + 1}) h(x_{n + 1}) \, \mu_{n + 1}(\rmd x_{n + 1}) = \int \nabla_\theta \ud{n ; \theta}(x_n, x_{n + 1}) h(x_{n + 1}) \, \mu_{n + 1}(\rmd x_{n + 1})
$$
and $\sup_{\theta \in \Theta} \int |\nabla_\theta \ud{n ; \theta}(x_n, x_{n + 1})| \, \mu_{n + 1}(\rmd x_{n + 1}) \leq c < \infty$ for some positive constant $c$, implying \hypref{assum:bias:bound} in the sense that for all $n \in \nset$, $h \in \bmf{\Xfd_{n + 1}}$, and $x_n \in \Xset$, 
$$
| \uk{n; \theta}h(x_n) - \uk{n; \theta'}h(x_n) | \leq c \| \theta - \theta' \| \| h \|_\infty
$$
(\emph{i.e.}, $\precpar = \| \theta - \theta' \|$ in this case). Finally, assume also that family satisfies \hypref{assum:strong:mixing} uniformly over the parameter space in the sense that for all $n \in \nset$, $(x_n, x_{n + 1}) \in \Xset^2$, and $\theta \in \Theta$, 
$
\udlow \leq \ud{n; \theta}(x_n, x_{n + 1}) \leq \udup. 
$  
Then Theorem~\ref{thm:bias:bound} provides a positive constant $d$ such that for all $n \in \nset$,  
$$
\big| \post{0:n ; \theta} h_n -  \post{0:n ; \theta'} h_n \big|  
\leq d n \| \theta - \theta' \| \sup_{k \in \nset} \| \addf{k} \|_\infty. 
$$
This extends previous results on the uniform continuity of the filter distribution (see, \eg, \cite{papavasiliou:2006,legland:oudjane:2004}) to smoothing of additive state functionals.  
\end{remark}
